---
title: "Bag of Little Bootstraps"
author: "Julia Webb"
date: "June 4, 2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bag of Little Bootstraps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
.math {
  font-size:small;
}
</style> 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Description
---
Bag of Little Bootstraps (blb) allows users to utilize bootstrapping to create regression models, specifically linear regression models or logistic regression models. Users are also given the option to use more than one CPU when fitting a model to the data. Note, if users wish to utilize this parallelization feature, they first must run `furrr::plan(multisession, worker = cl)` in the console (`cl` being the number of CPU's they wish to use.)

The library has two main functions, each of which come with a set of methods.

\newline

Functions and Methods 
---

##### Functions

1. `blblm(formula, data, m = 10, B = 5000, parallel = FALSE, cl = NULL)` : Utilize bootsrapping to fit a linear regression model, to the data as specified in the data argument. If `parallel = TRUE`, then parallelization is used. Then the user can define the number of desired number CPUs in the `cl` argument. However, as noted in the description, the user must run `furrr::plan(multisession, worker = cl)` in the console before using this parallelization feature. The arguments `B` and `m` are the number of bootstrapped samples, and the number of sugroups in which to split the data during the bootstrapping process.

    * Methods:
        * `print(x, ...)` : Prints the model in the form of $Y$ ~ $X_{1}$ + $X_{2}$
        
        * `sigma(object, confidence = FALSE, level = 0.95, ...)`: Returns estimate of $\sigma$, the standard deviation of the errors. If `confidence = TRUE`, then a confidence interval of is $\sigma$ included in the output. The default confidence level is 95%, and this can be changed using the `level` argument.
        
        * `coef(object, ...)`: Returns a numeric vector containing the coefficients of the parameters in the model. For example, in the model $Y$ ~ $B_{1}X_{1} + B_{2}X_{2}$, this function would return a numeric vector containing $B_{1}$ and $B_{2}$.
        
        
        * `confint(object, parm = NULL, level = 0.95, ...)`: By default, returns 95% confidence intervals for each parameter in the model (the predictor variables). User can request specific parameters using the `parm` argument, and can change the confidence level using the `level` argument. 
        
        
        * `predict(object, new_data, confidence = FALSE, level = 0.95, ...)`: Utilize the model to predict the response values of a new dataset. If `confidence = TRUE` a n $x$ 3 matrix is returned in which the first column is a the predicted value, and the second and third rows are the lower and upper endpoints of the corresponding 95% confidence interval for that prediction. The user can change the confidence level with the `level` argument. 



2. `blb_logreg(formula, data, m = 10, B = 5000, parallel = FALSE, cl = NULL)`: Utilize bootsrapping to fit a logistic regression model, to the data as specified in the data argument. If `parallel = TRUE`, then parallelization is used. Then the user can define the number of desired number CPUs in the `cl` argument. However, as noted in the description, the user must run `furrr::plan(multisession, worker = cl)` in the console before using this parallelization feature. The arguments `B` and `m` are the number of bootstrapped samples, and the number of sugroups in which to split the data during the bootstrapping process.

    * Methods:
    
        * `print(x, ...)`: Prints the model in the form of $Y$ ~ $X_{1}$ + $X_{2}$. Note, this is not the equation used to calculate the data, as we are using logistic regression. However, given that we have p parameters, the log-odds of a given observation is equal to $B_{1}X_{1} + B_{2}X_{2}$ + ... + $B_{p}X_{p}$. 
        
        * `coef(object, ...)`: Returns a numeric vector containing the coefficients of the parameters in the model. For example, if the log-odds were equal to $B_{1}X_{1} + B_{2}X_{2}$, this function would return a numeric vector containing $B_{1}$ and $B_{2}$.
        
        * `confint(object, parm = NULL, level = 0.95, ...)`: By default, returns 95% confidence intervals for each parameter in the model (the predictor variables). User can request specific parameters using the `parm` argument, and can change the confidence level using the `level` argument. 
        
        * `predict(object, new_data, p = FALSE,...)`: Utilize the model to predict the response values of a new dataset. If `p = TRUE`, then the function returns P(Y = 1|X = x), instead of binary {0,1} prediction.
            

\newline

Examples
---

```{r  setup}
library(blb)
```


##### Examples concerning `blblm()`
We will use the iris dataset from the dataset library. See following link for more information. https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/iris
```{r}
library(datasets)
data(iris)

train = iris[1:75,]
test = iris[76:150,]
```


Example 1: `blblm()` 

* default: 
```{r}
fit = blblm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width, data = train)
```

* change values of `m`, `B`, `parallel`, `cl`
```{r}
fit = blblm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width, data = train, m = 8, B = 100, parallel = TRUE, cl = 4)
```

\newline

Example 2: `print()` method:
```{r}
print(fit)
```

\newline

Example 3: `sigma()` method              

*default
```{r}
sigma(fit)
```

* including a 90% confidence interval
```{r}
sigma(fit, confidence = TRUE, level = 0.9)
```

\newline

Example 4: `coef()` method
```{r}
coef(fit)
```

\newline

Example 5: `confint()` 

* default
```{r}
confint(fit)
```

* specify parameters, change confidence level to 90%
```{r}
confint(fit, parm = "Sepal.Width", level = 0.90)



confint(fit, parm = c("Sepal.Width", "Petal.Width"), level = 0.90)
```

Example 6: `predict()` 

*default
```{r}
predicted_values = predict(fit, test)
head(predicted_values)
```

* `confidence = TRUE`, change confidence level to 90%
```{r}
predicted_values = predict(fit, test, confidence = TRUE, level = 0.90)
head(predicted_values)
```


##### Examples concerning `blb_logreg()`

We will use data from the `titanic` library. See following link for more information. https://www.rdocumentation.org/packages/titanic/versions/0.1.0

```{r}
library(titanic)

train = titanic_train[1:550,]
test = titanic_train[551:891,]
```


Example 1: `blb_logreg()`

* default
```{r}
fit_logreg = blb_logreg(Survived ~ Age + Sex + Pclass, data = train)
fit_logreg
```

* change values of `m`, `B`, `parallel`, `cl`
```{r}
fit_logreg = blb_logreg(Survived ~ Age + Sex + Pclass, data = train, m = 8, B = 100, parallel = TRUE, cl = 4)
fit_logreg
```

\newline

Example 2: `print()` method
```{r}
print(fit_logreg)
```

\newline
Example 2: `coef()` method

* Note: The `Sex` parameter can take on values in {Male, Female}. The blb_logreg() converts `Sex` to a binary variable , `Sexmale`, such that `Sexmale = 1` if `Sex = Male`, and `Sexmale = 0` if `Sex = Female`.
```{r}
coef(fit_logreg)
```

\newline

Example 3: `confint()` method  

* Note the use of `Sexmale` instead of original parameter `Sex`. See Example 2 for explanation.

* default output:
```{r}
confint(fit_logreg)
```

* specify parameters with `parm` argument, and change confidence level to 90%
```{r}
confint(fit_logreg, parm = c("Age", "Sexmale"), level = 0.9)



confint(fit_logreg, parm = c("Age"), level = 0.9)
```

Example 4: `predict()` method

* default output
```{r}
predicted_values = predict(fit_logreg, test)
head(predicted_values)
```

* `p = TRUE`
```{r}
predicted_values = predict(fit_logreg, test, p = TRUE)
head(predicted_values)
```
